---
title: "Project: MovieLens Recommendation System"
author: "Abraham Verde"
date: "March/4/2022"
output:
  pdf_document: default
  html_document: default
---

## INTRODUCTION 

For this project, I will be creating a movie recommendation system using the MovieLens dataset. The version of movielens included in the dslabs package is a small subset of a major dataset. A movie recomendation system pretends to predict the preference of a user. I'm going to use the rating values and movie genres in the dataset to make models for this prediction.

The data set used in this project is loaded using the code provided in the course. This code split the data into a train set named edx and test set named validation. It's important to notice, the test set (validation)is roughly 10% the train set (edx).


```{r error=FALSE, warning=FALSE}
##########################################################
# Create edx set, validation set (final hold-out test set)
##########################################################

# Note: this process could take a couple of minutes

if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(data.table)) install.packages("data.table", repos = "http://cran.us.r-project.org")

library(tidyverse)
library(caret)
library(data.table)

# MovieLens 10M dataset:
# https://grouplens.org/datasets/movielens/10m/
# http://files.grouplens.org/datasets/movielens/ml-10m.zip

dl <- tempfile()
download.file("http://files.grouplens.org/datasets/movielens/ml-10m.zip", dl)

ratings <- fread(text = gsub("::", "\t", readLines(unzip(dl, "ml-10M100K/ratings.dat"))),
                 col.names = c("userId", "movieId", "rating", "timestamp"))

movies <- str_split_fixed(readLines(unzip(dl, "ml-10M100K/movies.dat")), "\\::", 3)
colnames(movies) <- c("movieId", "title", "genres")

# if using R 3.6 or earlier:
movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(levels(movieId))[movieId],
                                           title = as.character(title),
                                           genres = as.character(genres))
# if using R 4.0 or later:
#movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(movieId),
#                                           title = as.character(title),
#                                           genres = as.character(genres))


movielens <- left_join(ratings, movies, by = "movieId")

# Validation set will be 10% of MovieLens data
set.seed(1, sample.kind="Rounding") # if using R 3.5 or earlier, use `set.seed(1)`
test_index <- createDataPartition(y = movielens$rating, times = 1, p = 0.1, list = FALSE)
edx <- movielens[-test_index,]
temp <- movielens[test_index,]

# Make sure userId and movieId in validation set are also in edx set
validation <- temp %>% 
  semi_join(edx, by = "movieId") %>%
  semi_join(edx, by = "userId")

# Add rows removed from validation set back into edx set
removed <- anti_join(temp, validation)
edx <- rbind(edx, removed)

rm(dl, ratings, movies, test_index, temp, movielens, removed)
```

# MY SOURCE CODE

## PRIMARY DATA EXPLORATION
Once I loaded the data set, is necessary to perform some exploration to get familiar with the data structures of the involved objects. In this first sight we can easily watch the edx and validation objects, both of them has six columns with the main information, such as: userId, movieId, rating, timestamp, title and movie geners.


```{r}
library(dplyr)
library(ggplot2)
library(tidyr)

#Exploration of main information
head(edx)
summary(edx)
```

#### Reviewing genre movie information
It's quite useful to get the genres information per movies. In the dataset this information comes non splited in the field genres. To get genres information per movies is necessary split it and create a new data frame with this information. Once I have the data frame already populated, I pretend to show this information in a bar graphs to watch the information in a confortable way.

```{r}
dataGR <- edx %>% separate_rows(genres, sep = "\\|")
GR_Rating <- dataGR %>% group_by(genres) %>% summarize(n=n()) %>% arrange(n)
##Convert genres to factor, to get order in graph
GR_Rating$genres <- factor(GR_Rating$genres, levels = GR_Rating$genres[order(GR_Rating$n)])
```

#### Graph genre movies
```{r}
graph_1<-ggplot(data=GR_Rating, aes(x=n, y=genres)) +
  geom_bar(stat="identity") + 
  ggtitle("Movie Quantity by Genre")
graph_1
```

#### Detect top 10 rating movies
Other information that is good to have is about top rated movies. In this case, I'm going to show only the first 10 highest ratings.

```{r}
top10 <- edx %>% group_by(title) %>% summarize(ratingMean=mean(rating)) %>% arrange(desc(ratingMean))

##Convert title to factor, to get order in graph
top10$title <- factor(top10$title, levels = top10$title[order(top10$ratingMean)])
top10[1:10,]
```

#### Graph top 10 rating movies
```{r}
graph_2<- top10[1:10, ] %>% ggplot(aes(x=ratingMean, y=title)) +
  geom_bar(stat="identity") + 
  ggtitle("Top 10 rating movies")
graph_2
```


#### Detect top 10 rating genre
As the previous graph, is good to know the top movie rating grouping by genres. In this case, I'm going to show only the first 10 highest rating per genres.

```{r}
top10_rating <- GR_Rating %>% group_by(genres) %>% arrange(desc(n))

##Convert genres to factor, to get order in graph
top10_rating$genres <- factor(top10_rating$genres, levels = top10_rating$genres[order(top10_rating$n)])
top10_rating[1:10, ]
```

#### Graph top 10 rating genre
```{r}
graph_3<- top10_rating[1:10, ] %>% ggplot(aes(x=n, y=genres)) +
  geom_bar(stat="identity") + 
  ggtitle("Top 10 rating genres")
graph_3
```


## ANALYSIS
The next step after the exploration data analysis, is to build some models that help us to get the prediction.


#### Define RMSE function
By definition The root mean square error (RMSE) allows us to measure how far predicted values are from observed values in a regression analysis. I build a function to get RMSE value.
```{r}
RMSE <- function(true_ratings, predicted_ratings){
  sqrt(mean((true_ratings - predicted_ratings)^2, na.rm = TRUE))
}
```

#### SPLIT THE DATASET EDX. 
It's not allowed to use validations set to train, so i'm going to split edx dataset into train and test set
```{r}
set.seed(1, sample.kind="Rounding") # if using R 3.5 or earlier, use `set.seed(1)`
tempSet <- createDataPartition(y = edx$rating, times = 1, p = 0.1, list = FALSE)
train_set <- edx[-tempSet,]
test_set <- edx[tempSet,]
```


#### First Approach - just rating average
The simplest approach is just the rating average, I'm going to take the average of the rating values from the train set.
```{r}
FirstApproach_Model <- mean(train_set$rating);
FirstApproach_Model
```

#### PREDICT - First Approach
Once I have the "model", get the RMSE values.

```{r}
FirstApproach_Predict<- RMSE(test_set$rating,FirstApproach_Model)
FirstApproach_Predict
```


#### Save in table the first approach
```{r}
rmse_results <- tibble(method = "First Approach - Average", RMSE = FirstApproach_Predict)
```

#### Show results in a table for comparison
```{r}
rmse_results
```


#### Second Approach - using movie effect
As second approach I'm going to add some features to the model to try to increase the accuracy of the prediction. In this case following the approach using in the course: https://rafalab.github.io/dsbook/large-datasets.html#modeling-movie-effects, I'm going to add movie effect to the model.

```{r}
avgs_of_movies <- train_set %>% group_by(movieId) %>% 
  summarize(bias_movies = mean(rating - FirstApproach_Model))
```


#### Second Approach - using movie effect
```{r}
SecondApproach_Model <- FirstApproach_Model + test_set %>% 
  left_join(avgs_of_movies, by='movieId') %>% pull(bias_movies)
```

#### PREDICT - Second Approach - using movie effect
In the next, I'm going to have the RMSE values for the second approach model.

```{r}
SecondApproach_Predict<- RMSE(test_set$rating,SecondApproach_Model) 
SecondApproach_Predict
```


#### Save in table the second approach
```{r error=FALSE, warning=FALSE}
rmse_results <- bind_rows(rmse_results, data_frame(method="Second Approach - Movie Effect",  RMSE = SecondApproach_Predict))
```

#### Show results in a table for comparison
```{r}
rmse_results
```


#### Third Approach - using user effect
Following the reference of the course: https://rafalab.github.io/dsbook/large-datasets.html#user-effects. It's possible increase the accuracy of the prediction if I consider to add one more effect to the model. In this case, I'm going to add the user effect.

Unlike the previous approach in this case I'm going to group by userID instead movieID. and the summary I have to add the bias that I got in the previous approach.
```{r}
avgs_of_users <- train_set %>% left_join(avgs_of_movies, by='movieId') %>% 
  group_by(userId) %>% 
  summarize(bias_user = mean(rating - FirstApproach_Model - bias_movies))
```

```{r}
#Here Build the model
ThirdApproach_Model <- test_set %>% left_join(avgs_of_movies, by='movieId') %>% 
  left_join(avgs_of_users, by='userId') %>% mutate(avgSum = (FirstApproach_Model + bias_movies + bias_user)) %>% 
  pull(avgSum)
```


#### PREDICT - Third Approach - using user effect
```{r}
ThirdApproach_Predict<- RMSE(test_set$rating,ThirdApproach_Model) 
ThirdApproach_Predict
```

#### Save in table the third approach
```{r}
rmse_results <- bind_rows(rmse_results, data_frame(method="Third Approach - Movie Effect/User Effect ",  RMSE = ThirdApproach_Predict))
```


#### Show results in a table for comparison
```{r}
rmse_results
```


## CHECK WITH VALIDATION SET
```{r}
validation_Model <- validation %>% 
  left_join(avgs_of_movies, by='movieId') %>% 
  left_join(avgs_of_users, by='userId') %>% 
  mutate(avgSum = FirstApproach_Model + bias_movies + bias_user) %>% 
  pull(avgSum)

validation_RMSE <- RMSE(validation$rating,validation_Model) 
validation_RMSE
```

#### Show results in a table for comparasion
```{r}
rmse_results <- bind_rows(rmse_results, data_frame(method="Validation",  RMSE = validation_RMSE))
```

## SHOW FINAL COMPARISON TABLE
We can easily detect that the third approach is way better than the first and second approach, in this case, I suggest to use the third approach to predict the recomendation system.
```{r}
rmse_results
```


## CONCLUSION
I already have developed three diferents approaches. The first one was only de rating average, this model or approach, give me a RMSE value higher than 1, in this case, I considered to add movie effect seeking to increase the accuracy of the prediction. Even the RMSE value actually improved, I considered to add one more effect (user effect) to try to improve as max as possible the accuracy. This final effect gives me the best RMSE I could achieve.


The second approach included the bs term that represent the average rating for each movie. With this approach, the RMSE value that I got was lower than 1.

The third approach included the bs term that represent the average rating for user. With this approach, the RMSE value is higher than the second approach.

In consequence, the third approach is the best and I recommended to use it.
